\LARGE \bf Expectation-Maximization based defense mechanism for distributed Model Predictive Control

### 2022-02-14 article/article.tex
change simulation, add data, Submitted to NecSys

\draft% show todos in red
% \final% give error if there is todos
% \comments% blank right column to make comments in pen & paper

@@

\begin{abstract}% Abstract of not more than 250 words.
  Controlling large scale systems sometimes requires decentralized computation.
  Communication among agents is crucial to achieve consensus and optimal global behavior.
  These negotiation mechanisms are sensitive to attacks on those exchanges.
  This paper proposes an algorithm based on Expectation Maximization to mitigate the effects of attacks in a resource allocation based distributed model predictive control.
  The performance is assessed through academical example of the temperature control of multiple rooms under input power constraints.
\end{abstract}

\begin{keyword}
  Model predictive and optimization-based control,
  Distributed control and estimation,
Large scale optimization problems
\end{keyword}

@@

\section{INTRODUCTION}
The information collected, processed, and transmitted within \cps{} impacts physical systems whose proper functioning is essential for society, like energy networks, industrial processes, water production plants, etc..
Attacks in those critical infrastructures, such as Stuxnet, increased awareness for cyber-security of \cps{}.

Given the complexity, scale and even geographic location of these structures, it is necessary to adopt control structures that are decentralized rather than monolithic.
For these systems, a set of controllers are designed to ensure the safe and efficient operation of the facilities collectively.

This operation is done cooperatively, where the controllers exchange information with each other.
This approach can be used, i.e., for the voltage regulation of electrical networks, or the production-consumption balance in these same networks, water distribution and many other applications.
Multi-agent systems~\citep{KantamneniEtAl2015} and \dmpc{}~\citep{MaestreEtAl2014} are standard approaches to design these controllers.

A few articles have focused on safety in \dmpc{} frameworks, presenting vulnerabilities of different frameworks, such as Lagrange-based decomposition~\citep{VelardeEtAl2017}, Jacobi-Gauss decomposition~\citep{ChanfreutEtAl2018} and primal decomposition~\citep{NogueiraEtAl2021}.

The authors in these works highlight the complexity of the problems. A modification of a local goal, the hacking of communication, or the failure of one of the agents are events that disrupt global behavior.
If one of these agents is attacked, it can result in the violent destruction of the corrupted element (perhaps the whole system), or more subtly, it can lead to deviant behavior that is more difficult to detect.
% Usually the attacks aim to de-optimize the global operation by benefiting/prejudicing some specific group of controllers, or to lead the system towards a state of malfunction/instability~\todo[use other]{}.

Some strategies are presented to mitigate the effects of such event.
For instance dismissing extreme values as in~\cite{VelardeEtAl2017}, or
using secure scenarios based on reliable historical data~\cite{MaestreEtAl2021}.

A previous work~\citep{NogueiraEtAl2021} proposed a safe algorithm based on data reconstruction to mitigate the effects of a \fdi{} in the communication between a coordinator and local constraint-free agents, which were bound by global equality constraints.

In this paper, we propose an extension of our approach by including local constraints and changing the global constraints.
This extension profoundly changes the complexity, as the exchanges between the
agents and the coordinator are no longer characterized by an affine functions
but by \pwa{} functions.
This leads not only to a combinatorial explosion but also to a parametric identification challenge.
To overcome this problem, we propose using a learning method based on the \EM{} algorithm~\citep{DempsterEtAl1977}, which allows us to estimate the corruption
mechanism and correct it if necessary.

This paper is organized as follows.
First, in Section~\ref{sec:PS}, we introduce the primal decomposition-based \dmpc{} and its vulnerabilities.
Then in Section~\ref{sec:TSM} we study the local problems and propose a detection and mitigation mechanism based on the Expectation Maximization algorithm.
Moreover, in Section~\ref{sec:App_dMPC}, we illustrate the algorithm with academic simulation and we assess its performance.
Finally, in Section~\ref{sec:CC}, we conclude with an outlook of future works.

\emph{Notation:}
In this paper, $\norm{\cdot}$, $\norm{\cdot}_{F}$, and $\norm{\vec{v}}_{Y}$ represent the $\ell_{2}$ norm, the Frobenius norm and the weighted norm $\norm{Y^{\frac{1}{2}}\vec{v}}$.
$\Proj^{\set{T}}(\cdot)$ is the Euclidean projection onto $\set{T}$.
% $\card{x}$ is the number of elements in $x$.
$\kron{}{}$ ($\hadamard{}{}$) is the Kronecker (Hadamard) product.
$\vectorize{A}$ vectorizes matrix $A$.
${n\mathbin{:}i\mathbin{:}j}$ is a row vector builder with elements $\{n,n+i,\dots,n+mi\}$, where ${m=\mathrm{truncate}(\frac{j-n}{i})}$, and ${n\mathbin{:}j}$ is equal to ${n\mathbin{:}1\mathbin{:}j}$.
$[A,B]$ ($[A;B]$) is horizontal (vertical) concatenation of matrices.
${\0_{m}=0_{m\times 1}}$ (${\1_{m}=1_{m\times 1}}$), where
$0_{m\times n}$ ($1_{m\times n}$) is a $0$ ($1$) filled \mbox{$m$-by-$n$} matrix.
$\indicator{x}$ is the indicator function returning $1$ if $x$ is true and $0$ otherwise.
$\preceq$ is the component-wise inequality for vectors and matrix inequality for matrices in ${\symmetric^{n}=\setbuild{X:\R^{n\times n}}{X\T=X}}$.
${\defpos^{n}=\setbuild{X:\symmetric^{n}}{X\succ 0}}$, ${\semidefpos^{n}=\setbuild{X:\symmetric^{n}}{X\succeq 0}}$.
$I_{c}$ is the identity matrix of size $c$.
$A^{\dagger}={{(A^{T}A)}^{-1}A^{T}}$.
$\determinant{A}$ is the determinant of $A$.
A vector $\vec{v}_{i}$, correspond to the $i$-th agent, and these vectors can be stacked in a vector $\vec{v}$.
$\random{a},\random{A},\randomvec{a}$ are random variables, scalar, matrix and vector respectively.
$\expectation{\random{x}}$ is the expected value of $\random{x}$, and $\probability{A \mid B}$ is the conditional probability of $A$ given condition $B$.
% $\pdf{\cdot}$ is a probability density function.
${\diag(A_{1},\dots,A_{N})}$ corresponds to a block diagonal matrix.

\section{Problem Statement}\label{sec:PS}
In this section we present an algorithm to decompose a \mpc{} problem based on the primal decomposition \citep{BoydEtAl2015}, and a vulnerability of the decomposition that can be exploited by a malicious agent.

@@

Each agent's input $\uik$ is constrained by
\begin{equation}
  \label{eq:u_local_constraints}
  \0_{n_{u}}\preceq\uik\preceq\vec{u}_{\max},
\end{equation}
with ${\umax:\R^{n_{u}}}$.
The agents are coupled by global input constraints with weighting matrices \todo{${\Gamma_{i}:\semidefpos^{n_{u}}}$}:
\begin{equation}
  \label{eq:global_constraint}
  \sum_{i\in\set{M}}\Gamma_{i}\uik\preceq\umax.
\end{equation}

@@

\end{matrix}
  $}
% \label{eq:dmpcModLOP}
\end{equation}
where ${Q_{i}:\semidefpos}$, ${R_{i}:\defpos}$, and $\vik$ is a control objective.
$\optglobobj$ denotes the optimal value of the objective function for the the optimal control sequences ${\optuiseq}$.
The problem in~\eqref{eq:GOP} is solved at each time $k$, and the ${\optuikk}$ are applied in their respective subsystem $i$, following a \rhs{}.

Solving~\eqref{eq:GOP} at each time $k$ can be quite challenging, since the prediction horizon and the number of agents can be large making the process computationally intensive. Another issue that can arise is the fact that the complete information \todo[verify]{(${\set{I}_{i}=\{A_{i},B_{i},C_{i},\vec{v}_{i}[k],Q_{i},R_{i}\}}$)} from all the subsystems is needed to solve the problem,
which can be viewed negatively in a confidentiality point of view.

So, we use the \emph{primal decomposition} method, sometimes called \emph{quantity decomposition} or \emph{resource allocation}, with wich we can profit from the parallelism, while avoiding the use of the complete information $\set{I}_{i}$.

@@

The technique consists of decomposing the monolithic \mpc{} in~\eqref{eq:GOP} into $M$ modified \mpc{} problems~\eqref{eq:DOP_local}, called \emph{local problems}, which can be solved in parallel by each subsystem, and a \emph{master problem}~\eqref{eq:DOP_master}, which is equivalent to the global problem and is solved by a coordinator.
\begin{subequations}
  \begin{equation}
    \left.
      \small
        J_{i}^{\star}(\thetaik)=
        \begin{matrix}
        \underset{\uiseq}{\mathrm{minimize}}&\obji\\
        \mathrm{subject~ to} &\eqref{eq:systems}\ \&\ \eqref{eq:u_local_constraints}\\
        &\Gamma_{i}\uik\preceq\qik:\dik\\
      \end{matrix}
    \right\}
    \small
    \begin{aligned}
      &\forall i\in\set{M}\\
      &\forall j\in\set{H}
    \end{aligned}
    \label{eq:DOP_local}
  \end{equation}

  \begin{equation}
    \small
    \begin{aligned}
      J^{\star}=
      \begin{matrix}
        \underset{\qiseq}{\mathrm{minimize}} &\sum_{i\in\set{M}} J^{\star}_i(\qik)\\
        \mathrm{subject~ to} &
        \left.
          \begin{aligned}
            \quad \sum_{i\in\set{M}}\qik\preceq\umax\\
            \quad \qik[k]\succeq\0_{n_{u}},       \forall i\in\set{M}
          \end{aligned}
        \right\}
        \begin{aligned}
          \forall j\in\set{H}
        \end{aligned}
      \end{matrix}
    \end{aligned}
    \label{eq:DOP_master}
  \end{equation}
\end{subequations}
The modified \mpc{} problems in~\eqref{eq:DOP_local} are created by exchanging the $\vec{u}_{\max}$ in~\eqref{eq:global_constraint} by vectors $\qik$, where which one corresponds to the quantity of the total resource $\vec{u}_{\max}$ that is allocated to agent $i$ in time $k$ for each prediction $j$, thus the names \emph{resource allocation} and \emph{quatity decomposition}. This new set of constraints have associated dual variables ${\dik\succeq\0_{n_{u}}}$. The sequence of this variables ${\qiseq}$ and ${\diseq}$ can be aggregated in vectors ${\thetaik=\kron{\1_{\predhorz}}{\qik}}$, and ${\lambdaik=\kron{\1_{\predhorz}}{\dik}}$.

@@

From~\cite{BoydEtAl2015}, it is known that $-\lambdaik$ are sub-gradients of objective $J[k]$.
Plugging it in $\vec{g}[k]\p$ we have
\begin{equation}
  \label{eq:thetaNegot}
\vec{\theta}[k]\pplusone=\Proj^{\set{S}}(\vec{\theta}[k]\p+\rho\p\vec{\lambda}[k]\p),
\end{equation}
\newcommand{\negotiation}{\emph{negotiation}}
which henceforth is referred as the \negotiation.

@@

The algorithm to find the optimal $\optthetai[k]$ can be summarized in algorithm~\ref{alg:quantityAlg}, and the exchange between coordinator and agents can be seen in Fig.~\ref{fig:dmpc_graph}.

\begin{algorithm2e}[h]
  \DontPrintSemicolon%
  $p:=0$\;
  Coordinator initializes $\vec{\theta}\p$ \;
  \Repeat{$\|\vec{\theta}^{(p)} -\vec{\theta}^{(p-1)}\|\leq\epsilon$}{
  Subsystems solve~\eqref{eq:DOP_local}, and send $\optlambdai (\thetai\p)$\;
  Coordinator updates allocations~\eqref{eq:thetaNegot}\;
  $p:=p+1$
}
 \caption{Quantity decomposition based \acrshort{dmpc}.}\label{alg:quantityAlg}
\end{algorithm2e}

One can observe that instead of using $\mathcal{I}_{i}$, each agent only sends $\lambdaik$.
An interpretation of $\lambdaik$ is the dissatisfaction of agent $i$ with the resource $\thetaik$ allocated for it, where ${\lambdaik=\0_{c}}$ means total satisfaction.
The coordinator trusts in the authenticity of the $\vec{\lambda}_{i}[k]$ received to update the allocations, but if false data is injected, the \negotiation{} can be driven by a malicious agent which can take advantage of this trust to favor itself, harm others or even destabilize the \negotiation{}, as shown in~\cite{NogueiraEtAl2021}.

@@

The \emph{local problems}~\eqref{eq:DOP_local} can be rewritten in a equivalent (same solution) constrained \qp{} form:
  \begin{subequations}\label{eq:quadratic_case}
    \begin{align}
      \label{eq:4}
      \underset{\Uik}{\mathrm{minimize}}&& \overbrace{\frac{1}{2} \Uik^T H_i\Uik+{\fik}^T\Uik}^{\textstyle J_{i}(\vec{\theta}_{i})}\\
      \mathrm{subject~ to}&&\bar{\Gamma}_{i}\Uik\preceq\vec{\theta}_{i}[k]:\vec{\lambda}_{i}[k]\label{eq:constraints}\\
                                        &&\Uik\succeq\0_{c},
    \end{align}
  \end{subequations}
where ${\Uik}$ stacks the input prediction sequences, and
${\bar{\Gamma}_{i}=\kron{I_{\predhorz}}{\Gamma_{i}}}$.
The exact values for $H_{i}$ and $\vec{f}_{i}[k]$ vary depending on the control objective $\vec{v}_{i}[k]$. The approach presented in this paper does not depend on the choice of the control objective.
The approach depends only on the fact that ${H_{i}:\semidefpos}$ does not depend on $k$, while $\fik$ does. Those facts are instrumental for the results in the following sections.

\no{
  We choose reference tracking just to illustrate the approach. With this objective, we have
${\vik=\wik-C_{i}\xik}$, where $\wik$ is the output reference in time $k$, which yields
\begin{equation}
\small\begin{matrix*}[l]
 H_i&=&\mathcal{D}_i^T\bar{Q}_i\mathcal{D}_i+\bar{R}_i\\
\fik&=&\mathcal{D}_i^T\bar{Q}_i(\mathcal{M}_i\xik-\Wik),
\end{matrix*}
\label{eq:matrices}
\end{equation}
where $\Wik$ stacks the setpoint prediction sequence, which is supposed to be constant during the predicted horizon ${\Wik=\kron{\1_{\predhorz}}{\vec{w}_{i}[k]}}$ (no prior information).
${\mathcal{M}_{i}}$  and ${\mathcal{D}_{i}}$ are the prediction matrices of the \mpc{}.
${\bar{Q}_{i}=\kron{I_{\predhorz}}{Q_{i}}}$, and ${\bar{R}_{i}=\kron{I_{\predhorz}}{R_{i}}}$.
}
\todo[make a better figure if time permits]{
\begin{figure}[t]
  \centering
  \begin{tikzpicture}[node distance=1.5cm]
    \tikzset{
      agent/.style={draw,circle,minimum width=1cm,},
      coord/.style={draw,circle,minimum width=1cm,very thick}
    }
    % \draw[thick,blue,rounded corners=10pt] (-1.7,1.2) rectangle (1.7,-1.7);
    % \node at (1.2,-1.3) {O};
    % \node (pi) at (-2.5,.5) {$\pi$};
    % \node (phi) at (2.5,-1.) {$\phi_{c}$};

    \node[agent] (a1) {a$_{1}$};
    \node (a2) [right of=a1]  {$\dots$};
    \node[agent] (am) [right of=a2] {a$_{M}$};
    \node[coord] (c) [above of=a2] {C};

    \draw[-latex] (a1) to [out=60,in=210] node [sloped,above] {$\vec{\lambda_{1}}[k]$} (c) ;
    \draw[-latex] (c) to [out=-120,in=30] node [sloped,below] {$\vec{\theta}_{1}[k]$} (a1);

    \draw[-latex] (am) to [out=150,in=-60]node [sloped,below] {$\vec{\lambda}_{M}[k]$} (c);
    \draw[-latex] (c) to [out=-30,in=120] node [sloped,above] {$\vec{\theta}_{M}[k]$} (am);

  \end{tikzpicture}
  \caption{Exchange between coordinator (C) and agents $a_{i}$.}\label{fig:dmpc_graph}
\end{figure}
}

@@

The halfspaces defined by the set of pairs $(G^{n}[k],b^{n}[k])$ represent the combinations of active constraints~\eqref{eq:constraints} for a given time $k$, which vary w.r.t. $\vec{v}_{i}[k]$.

The $P_{i}^{n}$ are constructed using $H_{i}$ and $\bar{\Gamma}_{i}$; and $\vec{s}_{i}^{j}$ are constructed using $H_{i}$, $\fik$ and $\bar{\Gamma}_{i}$. The specific construction depends on the active constraints in~\eqref{eq:constraints}.
For example, if all constraints are active, we have ${P_{i}^{\text{ac}}={(\bar{\Gamma}_{i}H_{i}^{-1}\bar{\Gamma}_{i}\T)}^{-1}}$ and ${\vec{s}_{i}^{\text{ac}}[k]=P_{i}^{1}\bar{\Gamma}_{i}H_{i}^{-1}\fik}$.
And if all constraints are inactive, we have ${P_{i}^{\text{in}}=0_{c\times c}}$ and ${\vec{s}_{i}^{\text{in}}=\0_{c}}$.

>>>>

\draft% show todos in red
\final% give error if there is todos
% \comments% blank right column to make comments in pen & paper

@@

\begin{abstract}% Abstract of not more than 250 words.
  Controlling large-scale systems sometimes requires decentralized computation.
  Communication among agents is crucial to achieving consensus and optimal global behavior.
  These negotiation mechanisms are sensitive to attacks on those exchanges.
  This paper proposes an algorithm based on Expectation Maximization to mitigate the effects of attacks in a resource allocation based distributed model predictive control.
  The performance is assessed through an academic example of the temperature control of multiple rooms under input power constraints.
\end{abstract}

\begin{keyword}
%   Model predictive and optimization-based control,
%   Distributed control and estimation,
% Large scale optimization problems
Distributed constrained control and MPC;
Cyber security in networked control systems;
Decentralized Control and Large-Scale Systems
\end{keyword}

@@

\section{INTRODUCTION}
The information collected, processed, and transmitted within \cps{} impacts physical systems whose proper functioning is essential for society, like energy networks, industrial processes, water distribution, and others.
Attacks in those critical infrastructures, such as Stuxnet, increased awareness for cyber-security of \cps{}.

Given these structures' complexity, scale, and geographic distribution, it is necessary to adopt decentralized control structures rather than monolithic ones.
For these systems, a set of controllers are designed to collectively ensure the safe and efficient operation of the facilities.

This operation is done cooperatively, where the controllers exchange information.
This approach can be used, i.e., for the voltage regulation of electrical networks, the production-consumption balance in these same networks, water distribution, and many other applications.
Multi-agent systems~\citep{KantamneniEtAl2015} and \dmpc{}~\citep{MaestreEtAl2014} are standard approaches to design these controllers.

A few articles have focused on safety in \dmpc{} frameworks, presenting vulnerabilities of different frameworks, such as Lagrange-based decomposition~\citep{VelardeEtAl2017}, Jacobi-Gauss decomposition~\citep{ChanfreutEtAl2018}, and primal decomposition~\citep{NogueiraEtAl2021}.

The authors in these works highlight the complexity of the problems. A modification of a local goal, the hacking of communication, or the failure of one of the agents are events that disrupt global behavior.
If one of these agents is attacked, it can result in the violent destruction of the corrupted element (perhaps the whole system), or more subtly, it can lead to a deviant behavior that is more difficult to detect.
% Usually the attacks aim to de-optimize the global operation by benefiting/prejudicing some specific group of controllers, or to lead the system towards a state of malfunction/instability~\todo[use other]{}.

Some strategies are presented to mitigate the effects of such an event.
For instance, dismissing extreme values as in~\cite{VelardeEtAl2017} or
using secure scenarios based on reliable historical data~\cite{MaestreEtAl2021}.

\cite{NogueiraEtAl2021} proposed a safe algorithm based on data reconstruction to mitigate the effects of a \fdi{} in the communication between a coordinator and local constraint-free agents, which were bound by global equality constraints.

In this paper, we propose an extension of their approach by including local constraints and changing the global constraints.
This extension profoundly changes the complexity, as the exchanges between the agents and the coordinator are no longer characterized by affine functions but by \pwa{} functions.
This fact leads not only to a combinatorial explosion but also to a parametric identification challenge.
To overcome this problem, we propose using a learning method based on the \EM{} algorithm~\citep{DempsterEtAl1977}, which allows us to estimate the corruption mechanism and correct it if necessary.

This paper is organized as follows.
First, Section~\ref{sec:PS} introduces the primal decomposition-based \dmpc{} and its vulnerabilities.
Then in Section~\ref{sec:TSM}, we study the local problems and propose a detection and mitigation mechanism based on the Expectation Maximization algorithm.
Moreover, in Section~\ref{sec:App_dMPC}, we illustrate the algorithm with an academic simulation and assess its performance.
Finally, in Section~\ref{sec:CC}, we conclude with an outlook of future works.

\emph{Notation:}
In this paper, $\norm{\cdot}$, $\norm{\cdot}_{F}$, and ${\norm{\vec{v}}_{Y}=\norm{Y^{\frac{1}{2}}\vec{v}}}$ represent the $\ell_{2}$, Frobenius and weighted norms.
$\Proj^{\set{T}}(\cdot)$ is the Euclidean projection onto $\set{T}$.
% $\card{x}$ is the number of elements in $x$.
$\kron{}{}$ ($\hadamard{}{}$) is the Kronecker (Hadamard) product.
$\vectorize{A}$ vectorizes matrix $A$.
${n\mathbin{:}i\mathbin{:}j}$ is a row vector builder with elements $\{n,n+i,\dots,n+mi\}$, where ${m=\mathrm{truncate}(\frac{j-n}{i})}$, and ${n\mathbin{:}j}$ means ${i=1}$.
% $[A,B]$ ($[A;B]$) is horizontal (vertical) concatenation of matrices.
${\0_{m}=0_{m\times 1}}$ (${\1_{m}=1_{m\times 1}}$), where
$0_{m\times n}$ ($1_{m\times n}$) is a $0$ ($1$) filled \mbox{$m$-by-$n$} matrix.
$\indicator{x}$ is the indicator function returning $1$ if $x$ is true and $0$ otherwise.
$\symmetric^{n}$, and $\defpos^{n}$ ($\semidefpos^{n}$) are symmetric and positive (semi-)definite matrices of size $n$.
${A^{\dagger}={{(A^{T}A)}^{-1}A^{T}}}$.
% $\determinant{A}$ is the determinant of $A$.
A vector $\vec{v}_{i}$, correspond to the $i$-th agent, and these vectors can be stacked in a vector $\vec{v}$.
% $\random{a},\random{A},\randomvec{a}$ are random variables, scalar, matrix and vector respectively.
$\expectation{\random{x}}$ is the expected value of random variable $\random{x}$, and $\probability{A \mid B}$ is the conditional probability of $A$ given condition $B$.
% $\pdf{\cdot}$ is a probability density function.
${\diag(A_{1},\dots,A_{N})}$ corresponds to a block diagonal matrix.

\section{Problem Statement}\label{sec:PS}
In this section, we present a \dmpc{} algorithm based on the primal decomposition \citep{BoydEtAl2015} and a vulnerability of the decomposition, which can be exploited.

@@

Each agent's input $\uik$ is constrained by
\begin{equation}
  \label{eq:u_local_constraints}
  \0_{n_{u}}\preceq\uik\preceq\vec{u}_{\max},
\end{equation}
with ${\umax:\R^{n_{u}}}$.
The agents are coupled by global input constraints with weighting matrices ${\Gamma_{i}:\semidefpos^{n_{u}}}$:
\begin{equation}
  \label{eq:global_constraint}
  \sum_{i\in\set{M}}\Gamma_{i}\uik\preceq\umax.
\end{equation}

@@

\end{matrix}
  $}
% \label{eq:dmpcModLOP}
\end{equation}
where ${Q_{i}:\semidefpos}$, ${R_{i}:\defpos}$, and $\vik$ is a control objective.
$\optglobobj$ denotes the optimal value of the objective function for the optimal control sequences ${\optuiseq}$.
The problem in~\eqref{eq:GOP} is solved at each time $k$, and the ${\optuikk}$ are applied in their respective subsystem $i$, following a \rhs{}.

Since the prediction horizon and the number of agents can be large, solving~\eqref{eq:GOP} at each time $k$ can be rather challenging due to computational costs.
Another issue that can arise is the fact that the complete information ${\set{I}_i=\{A_i,B_i,C_i,Q_i,R_i,\Gamma_i,\vec{v}_i\}}$ from all the subsystems is needed to solve the problem,
which can be viewed negatively from a confidentiality point of view.

So, we use the \emph{primal decomposition} method, sometimes called \emph{quantity decomposition} or \emph{resource allocation}, with which we can profit from the parallelism, while avoiding the use of the complete information $\set{I}_{i}$.

@@

The technique decomposes the monolithic \mpc{} in~\eqref{eq:GOP} into $M$ modified \mpc{} problems~\eqref{eq:DOP_local} (solvable in parallel by each subsystem) called \emph{local problems}, and a \emph{master problem}~\eqref{eq:DOP_master}, which is equivalent to the global problem and is solved by a coordinator.
\begin{subequations}
  \begin{equation}
    \left.
      \small
        J_{i}^{\star}(\thetaik)=
        \begin{matrix}
        \underset{\uiseq}{\mathrm{minimize}}&\obji\\
        \mathrm{subject~ to} &\eqref{eq:systems}\ \&\ \eqref{eq:u_local_constraints}\\
        &\Gamma_{i}\uik\preceq\qik:\dik\\
      \end{matrix}
    \right\}
    \small
    \begin{aligned}
      &\forall i\in\set{M}\\
      &\forall j\in\set{H}
    \end{aligned}
    \label{eq:DOP_local}
  \end{equation}

  \begin{equation}
    \small
    \begin{aligned}
      J^{\star}=
      \begin{matrix}
        \underset{\qiseq}{\mathrm{minimize}} &\sum_{i\in\set{M}} J^{\star}_i(\qik)\\
        \mathrm{subject~ to} &
        \left.
          \begin{aligned}
            \quad \sum_{i\in\set{M}}\qik\preceq\umax\\
            \quad \qik[k]\succeq\0_{n_{u}},       \forall i\in\set{M}
          \end{aligned}
        \right\}
        \begin{aligned}
          \forall j\in\set{H}
        \end{aligned}
      \end{matrix}
    \end{aligned}
    \label{eq:DOP_master}
  \end{equation}
\end{subequations}
The modified \mpc{} problems~\eqref{eq:DOP_local} are created by exchanging the $\vec{u}_{\max}$ in~\eqref{eq:global_constraint} by vectors $\qik$, which correspond to the quantity of the total resource $\vec{u}_{\max}$ allocated to agent $i$ in time $k$ for each prediction $j$, thus the names \emph{resource allocation} and \emph{quantity decomposition}.
This new set of constraints have associated dual variables ${\dik\succeq\0_{n_{u}}}$.
The sequences ${\qiseq}$ and ${\diseq}$ can be aggregated in vectors ${\thetaik=\kron{\1_{\predhorz}}{\qik}}$, and ${\lambdaik=\kron{\1_{\predhorz}}{\dik}}$.

@@

From~\cite{BoydEtAl2015}, it is known that $-\lambdaik$ are sub-gradients of objective $J[k]$.
Plugging it in $\vec{g}[k]\p$ we have
\begin{equation}
  \label{eq:thetaNegot}
\vec{\theta}[k]\pplusone=\Proj^{\set{S}}(\vec{\theta}[k]\p+\rho\p\vec{\lambda}[k]\p),
\end{equation}
\newcommand{\negotiation}{\emph{negotiation}}
which henceforth is referred to as the \negotiation.

@@

The algorithm to find the optimal $\optthetai[k]$ can be summarized in Algorithm~\ref{alg:quantityAlg}, and the exchange between coordinator and agents can be seen in Fig.~\ref{fig:dmpc_graph}.

\begin{figure}[h]
  \centering
    \resizebox{.47\columnwidth}{!}{
          \begin{tikzpicture}[font=\small,thick,node distance=.0cm and .5cm,
          mpcSmall/.style={
            rectangle,
            align=center,
            fill=supelecRed!10,
            minimum width=1.5cm,
            minimum height=1cm},
          coordinator/.style={
            rectangle,
            align=center,
            fill=mpc_grey,
            minimum height=3.4cm,
            minimum width=1.5cm,
          },
          superv/.style={
            rectangle,
            minimum height=1.4cm,
            minimum width=.3cm,
            fill=mpc_yellow,
          },
          negotiation/.style={
            rectangle,
            minimum height=3.0cm,
            minimum width=.45cm,
            fill=mpc_green,
          },
          ]

          \node[draw,
          mpcSmall,
          fill=mpc_blue,
          ] (block1) {MPC 1};

          \node[draw,
          mpcSmall,
          fill=none,
          draw=white,
          below=of block1,
          ] (mult) {\vdots};

          \node[draw,
          mpcSmall,
          fill=mpc_blue,
          below=of mult,
          ] (blockM) {MPC M};

          \node[draw,
          coordinator,
          right=2cm of mult,
          ] (coordinator) {};
          \node at ($(coordinator)+(.25,0)$) {\rotatebox{90}{Coordinator}};

          \node[draw,negotiation] (negot) at ({$(coordinator.center)+(0,0)$} -|{$(coordinator.west)+(0.25,)$}) {};
          \node  at (negot) {\rotatebox{90}{Negotiation}};

          \draw[latex-,thick] (block1.east)+(0,.3) -- ( {$(block1.east)+(0,.3)$} -| coordinator.west) node [above,midway] {$\tilde{\vec{\lambda}}_{1}[k]$};
          \draw[-latex,thick] (block1.east)+(0,-.3) -- ( {$(block1.east)+(0,-.3)$} -| coordinator.west) node [below,midway] {$\vec{\theta}_{1}[k]$};

          \draw[latex-,thick] (blockM.east)+(0,.3) -- ( {$(blockM.east)+(0,.3)$} -| coordinator.west) node [above,midway] {$\tilde{\vec{\lambda}}_{M}[k]$};
          \draw[-latex,thick] (blockM.east)+(0,-.3) -- ( {$(blockM.east)+(0,-.3)$} -| coordinator.west) node [below,midway] {$\vec{\theta_{M}}[k]$};

        \end{tikzpicture}
        }
  \caption{Exchange between coordinator and agents.}\label{fig:dmpc_graph}
\end{figure}

\begin{algorithm2e}[h]
  \DontPrintSemicolon%
  $p:=0$\;
  Coordinator initializes $\vec{\theta}\p$ \;
  \Repeat{$\|\vec{\theta}^{(p)} -\vec{\theta}^{(p-1)}\|\leq\epsilon$}{
  Subsystems solve~\eqref{eq:DOP_local}, and send $\optlambdai (\thetai\p)$\;
  Coordinator updates allocations~\eqref{eq:thetaNegot}\;
  $p:=p+1$
}
 \caption{Quantity decomposition-based \acrshort{dmpc}.}\label{alg:quantityAlg}
\end{algorithm2e}

One can observe that each agent only sends $\lambdaik$ instead of using $\mathcal{I}_{i}$.
An interpretation of $\lambdaik$ is the dissatisfaction of agent $i$ with the resource $\thetaik$ allocated for it, where ${\lambdaik=\0_{c}}$ means total satisfaction.
The coordinator trusts in the authenticity of the $\vec{\lambda}_{i}[k]$ received to update the allocations. However, if false data is injected, the \negotiation{} can be driven by a malicious agent, taking advantage of this trust to favor itself, harm others or even destabilize the \negotiation{}, as shown in~\cite{NogueiraEtAl2021}.

@@

The \emph{local problems}~\eqref{eq:DOP_local} can be rewritten in a equivalent (same solution) constrained \qp{} form:
  \begin{subequations}\label{eq:quadratic_case}
    \begin{align}
      \label{eq:4}
      \underset{\Uik}{\mathrm{minimize}}&&\small J_{i}(\vec{\theta}_{i})=&\frac{1}{2} \Uik^T H_i\Uik +{\fik}^T\Uik\\
      \mathrm{subject~ to}&&&\bar{\Gamma}_{i}\Uik\preceq\vec{\theta}_{i}[k]:\vec{\lambda}_{i}[k]\label{eq:constraints}\\
                                        &&&\Uik\succeq\0_{c},
    \end{align}
  \end{subequations}
where ${\Uik}$ stacks the input prediction sequences, and
${\bar{\Gamma}_{i}=\kron{I_{\predhorz}}{\Gamma_{i}}}$.
The values for $H_{i}$ and $\vec{f}_{i}[k]$ vary depending on the control objective $\vec{v}_{i}[k]$. The approach presented in this paper does not depend on the choice of the control objective.
The approach depends only on the fact that ${H_{i}:\semidefpos}$ does not depend on $k$, while $\fik$ does. Those facts are instrumental for the results in the following sections.

\no{
  We choose reference tracking to illustrate the approach. With this objective, we have
${\vik=\wik-C_{i}\xik}$, where $\wik$ is the output reference in time $k$, which yields
\begin{equation}
\small\begin{matrix*}[l]
 H_i&=&\mathcal{D}_i^T\bar{Q}_i\mathcal{D}_i+\bar{R}_i\\
\fik&=&\mathcal{D}_i^T\bar{Q}_i(\mathcal{M}_i\xik-\Wik),
\end{matrix*}
\label{eq:matrices}
\end{equation}
where $\Wik$ stacks the setpoint prediction sequence, which is supposed to be constant during the predicted horizon ${\Wik=\kron{\1_{\predhorz}}{\vec{w}_{i}[k]}}$ (no prior information).
${\mathcal{M}_{i}}$  and ${\mathcal{D}_{i}}$ are the prediction matrices of the \mpc{}.
${\bar{Q}_{i}=\kron{I_{\predhorz}}{Q_{i}}}$, and ${\bar{R}_{i}=\kron{I_{\predhorz}}{R_{i}}}$.
}

@@

The halfspaces defined by the pairs $(G^{n}[k],b^{n}[k])$ represent the combinations of active constraints~\eqref{eq:constraints} for a given time $k$, which vary w.r.t. $\vec{v}_{i}[k]$.

The $P_{i}^{n}$ are constructed using $H_{i}$ and $\bar{\Gamma}_{i}$; and $\vec{s}_{i}^{j}$ are constructed using $H_{i}$, $\fik$ and $\bar{\Gamma}_{i}$. The specific construction depends on the active constraints in~\eqref{eq:constraints}.
For example, if all constraints are active, we have ${P_{i}^{\text{ac}}={(\bar{\Gamma}_{i}H_{i}^{-1}\bar{\Gamma}_{i}\T)}^{-1}}$ and ${\vec{s}_{i}^{\text{ac}}[k]=P_{i}^{1}\bar{\Gamma}_{i}H_{i}^{-1}\fik}$.
Futhermore, if all constraints are inactive, we have ${P_{i}^{\text{in}}=0_{c\times c}}$ and ${\vec{s}_{i}^{\text{in}}=\0_{c}}$.

% \begin{figure}[t]
%   \centering
%           \begin{tikzpicture}[font=\small,thick,node distance=.0cm and .5cm,
%           mpcSmall/.style={
%             rectangle,
%             align=center,
%             fill=supelecRed!10,
%             minimum width=2.cm,
%             minimum height=1cm},
%           coordinator/.style={
%             rectangle,
%             align=center,
%             fill=mpc_grey,
%             minimum width=1cm,
%             minimum height=3.5cm
%           } ]

%           \node[draw,
%           mpcSmall,
%           fill=mpc_blue,
%           ] (block1) {MPC 1};

%           \node[draw,
%           mpcSmall,
%           fill=none,
%           draw=white,
%           below=of block1,
%           ] (mult) {\vdots};

%           \node[draw,
%           mpcSmall,
%           fill=mpc_blue,
%           below=of mult,
%           ] (blockM) {MPC M};

%           \node[draw,
%           coordinator,
%           right=2cm of mult,
%           ] (coordinator) {\rotatebox{90}{Coordinator}};
